### 2023-04-13 peer_review/model-free-reinforcement-learning_kinnunen_review.tex

\documentclass[a4paper,11pt]{article}

\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{varwidth,xcolor}
\usepackage{graphicx}
\usepackage{amssymb}

\newcommand{\veps}{\varepsilon}
\newcommand{\ee}{{\rm e}\hspace{1pt}}

\begin{document}
\pagestyle{empty}

  \begin{center}
{\large {\bf 
Review on a report \\
``Title of the report'' \\
by Report Author} \\

\bigskip

Written by:

 } 
\end{center}

Text text text text text text text text text text text text text text text text text text text text text text.
Text text text text text text text text text text text text text text text text text text text text text text.
Text text text text text text text text text text text text text text text text text text text text text text.
Text text text text text text text text text text text text text text text text text text text text text text.
Text text text text text text text text text text text text text text text text text text text text text text.
Text text text text text text text text text text text text text text text text text text text text text text.
Text text text text text text text text text text text text text text text text text text text text text text.

However, I do have some suggestions for corrections,
described in more detail below:

\begin{itemize}

	\item Text text text text text text text text text text text text text text text text text text text text text text.
	\item Text text text text text text text text text text text text text text text text text text text text text text.
	\item Text text text text text text text text text text text text text text text text text text text text text text.
	\item Text text text text text text text text text text text text text text text text text text text text text text.
	\item Text text text text text text text text text text text text text text text text text text text text text text.
	\item Text text text text text text text text text text text text text text text text text text text text text text.
	\item Text text text text text text text text text text text text text text text text text text text text text text.
	\item Text text text text text text text text text text text text text text text text text text text text text text.
	\item Text text text text text text text text text text text text text text text text text text text text text text.
	\item Text text text text text text text text text text text text text text text text text text text text text text.
	\item Text text text text text text text text text text text text text text text text text text text text text text.
	\item Text text text text text text text text text text text text text text text text text text text text text text.
	\item Text text text text text text text text text text text text text text text text text text text text text text.

\end{itemize}

\end{document}